---
title: "Preregistration Analysis"
author: "Natalie Biderman"
date: "11/21/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

Based on analysis of Exp1.

# Setup 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r general setup, echo=FALSE, results='hide'}

rm(list=ls(all=TRUE)) 

# If packages are not installed, install. Then, load libraries. 
list_of_packages <- c("ggplot2", "Rmisc", "cowplot", "reshape2", "gridExtra", "arm", "mosaic", "stringr", "tidyr", "dplyr", "bayesplot", "rstanarm", "arm")
new_packages <- list_of_packages[!(list_of_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)
lapply(list_of_packages, require, character.only = TRUE)

# Load functions
source("../Analysis/code/Functions/plotting.R")

# Figure parameters
fig_size = c(10,10)
fig_type = "eps" # or eps
Save_plots <- 1
```

```{r load data and model coefs}

# Load data 
load("../Analysis/data/Clean_data_lists/clean_data_Exp1.RData")
load("../Analysis/data/Clean_data_lists/clean_data_Exp2.RData")
load("../Analysis/data/Clean_data_lists/clean_data_Exp3.RData")
load("../Analysis/data/Clean_data_lists/clean_data_Exp4.RData")
load("../Analysis/data/Clean_data_lists/clean_data_Exp5.RData")
load("../Analysis/data/Clean_data_lists/clean_data_Exp6.RData")


# Load model coefs
load("../Analysis/data/Models/Choice_Ratings_models/Coef_lists/coef_list_Exp1.RData")
load("../Analysis/data/Models/Choice_Ratings_models/Coef_lists/coef_list_Exp2.RData")
load("../Analysis/data/Models/Choice_Ratings_models/Coef_lists/coef_list_Exp3.RData")
load("../Analysis/data/Models/Choice_Ratings_models/Coef_lists/coef_list_Exp4.RData")
load("../Analysis/data/Models/Choice_Ratings_models/Coef_lists/coef_list_Exp5.RData")
load("../Analysis/data/Models/Choice_Ratings_models/Coef_lists/coef_list_Exp6.RData")

```

```{r bind experiments into data frames}

bind_experiments <- function(phase_types, exps){
  all_dfs_list <- list()
  for (p in 1:length(phase_types)){
    data <- c()
    for (i in 1:length(exps)){
      curr_data <- eval(parse(text = sprintf("clean_data_%s$%s",exps[i],phase_types[p])))
      data <- bind_rows(data,curr_data)
    }
    all_dfs_list[[p]] <- data
  }
  names(all_dfs_list) <- phase_types
  return(all_dfs_list)
}

exps_to_be_used <- c("Exp1", "Exp2" ,"Exp3", "Exp4", "Exp5", "Exp6")
phase_types <- names(clean_data_Exp2)[-c(1:2)]
all_exps_list <- bind_experiments(phase_types, exps_to_be_used)
n_exps <- length(unique(all_exps_list$final_decisions$Exp))

```

## Power analysis for Experiment 1 (simple logistic regressions)

```{r}

library(lme4)
library(arm)
library(pwr)
library(lsr)

# === simple logistic regressions to predict decision bias ===

final_decisions_exp1 <- subset(all_exps_list$final_decisions, Exp=="Exp1" & !is.na(left_chosen)) 
# run logistic fits for each subject and detect their unchosen intercept
subs <- unique(final_decisions_exp1$PID)
subs_coefs_bias <- data.frame()
for (i in 1:length(subs)){
  sub_data <- subset(final_decisions_exp1, PID == subs[i])
  m_sub <- glm(data = sub_data, 
                higher_outcome_chosen ~ chosen_trial_centered * norm_drate_by_outcome, 
                family = binomial(link = "logit"))
  subs_coefs_bias[i,1] <- subs[i]
  subs_coefs_bias[i,c(2:5)] <- m_sub$coefficients
}
colnames(subs_coefs_bias) <- c("PID","(Intercept)","chosen_trial_centered","norm_drate_by_outcome","chosen_trial_centered:norm_drate_by_outcome")
subs_coefs_bias <- mutate(subs_coefs_bias, unchosen_bias_intercept = `(Intercept)` - chosen_trial_centered)
# compute power
compute_power <- function(desired_power,desired_sig_level,data,coef,null_point){
  t_stats <- data %>%
  dplyr::summarize(t_value = as.numeric(t.test(!!sym(coef), rep(null_point, n()), paired=TRUE)["statistic"]),
            p_value_round = round(as.numeric(t.test(!!sym(coef), rep(null_point, n()), paired=TRUE)["p.value"]),3),
            p_value = as.numeric(t.test(!!sym(coef), rep(null_point, n()), paired=TRUE)["p.value"]),
            cohens_d = cohensD(x=!!sym(coef),y=rep(null_point, n()),method="paired"), 
            power = as.numeric(pwr.t.test(n = n(), d = cohens_d, sig.level = desired_sig_level, 
                               type = c("paired"))["power"]),
            desired_sample = as.numeric(pwr.t.test(power = desired_power, d = cohens_d, 
                                        sig.level = desired_sig_level, type = c("paired"))["n"]))
  return(t_stats)
}
compute_power(0.99, 0.05, subs_coefs_bias, "unchosen_bias_intercept",0)

# === correlation between bias effect and pairs memory ===

bias_effect <- subset(all_exps_list$final_decisions, Exp=="Exp1") %>%
  group_by(PID,chosen_trial_name) %>%
  dplyr::summarize(p_gain = mean(higher_outcome_chosen, na.rm=1)) %>%
  spread(chosen_trial_name, p_gain) %>%
  mutate(bias_effect = Chosen-Unchosen)
memory_effect <- subset(all_exps_list$memory, Exp=="Exp1") %>%
  group_by(PID) %>%
  dplyr::summarise(pairs_acc = mean(pair_acc, na.rm=1))
bias_memory <- merge(bias_effect, memory_effect, by="PID")

r_stats <- bias_memory %>%
  dplyr::summarize(
    r_value = as.numeric(cor_test(bias_effect, pairs_acc)[[4]]),
    p_value = as.numeric(cor_test(bias_effect, pairs_acc)[[3]]),
    power = as.numeric(pwr.r.test(n = n(), r = r_value, sig.level = 0.05)["power"]),
    desired_sample = as.numeric(pwr.r.test(power = 0.99, r = r_value, sig.level = 0.05)["n"]))
print(r_stats)

```

## assess power for exp4 - memory manipulation

```{r}

final_decisions_exp4 <- subset(all_exps_list$final_decisions, Exp=="Exp4" & !is.na(left_chosen)) %>%
  mutate(repeat_cond_centered = ifelse(repeat_cond==1,1,-1),
         norm_drate_by_outcome = ifelse(gain_left, norm_rate_left-norm_rate_right,
                                        norm_rate_right-norm_rate_left))
m_bias_exp4 <- glmer(data = final_decisions_exp4,
                   higher_outcome_chosen ~ chosen_trial_centered * norm_drate_by_outcome * repeat_cond_centered +
                     (chosen_trial_centered * norm_drate_by_outcome * repeat_cond_centered| PID),
                   family = binomial(link = "logit"),
                   control=glmerControl(optCtrl=list(maxfun=2e4)))

# model per subject
subs <- unique(final_decisions_exp4$PID)
subs_coefs_bias_exp4 <- data.frame()
for (i in 1:length(subs)){
  sub_data <- subset(final_decisions_exp4, PID == subs[i])
  m_sub <- glm(data = sub_data, 
                higher_outcome_chosen ~ chosen_trial_centered * norm_drate_by_outcome * repeat_cond_centered,
                family = binomial(link = "logit"))
  subs_coefs_bias_exp4[i,1] <- subs[i]
  subs_coefs_bias_exp4[i,c(2:9)] <- m_sub$coefficients
}
colnames(subs_coefs_bias_exp4) <- c("PID",rownames(as.data.frame(m_sub$coefficients)))
subs_coefs_bias_exp4 <- mutate(subs_coefs_bias_exp4, 
                               unchosen_intercept_repeat = `(Intercept)` - chosen_trial_centered +
                                 repeat_cond_centered - `chosen_trial_centered:repeat_cond_centered`,
                               unchosen_intercept_no_repeat = `(Intercept)` - chosen_trial_centered -
                                 repeat_cond_centered + `chosen_trial_centered:repeat_cond_centered`,
                               unchosen_intercept_cond = 2*(repeat_cond_centered -
                                 `chosen_trial_centered:repeat_cond_centered`))

compute_power(0.80, 0.05, subs_coefs_bias_exp4, "unchosen_intercept_repeat",0)
compute_power(0.80, 0.05, subs_coefs_bias_exp4, "unchosen_intercept_no_repeat",0)
compute_power(0.80, 0.05, subs_coefs_bias_exp4, "unchosen_intercept_cond",0)

bias_exp4 <- final_decisions_exp4 %>% group_by(PID, chosen_trial_name, repeat_cond_name) %>%
  dplyr::summarize(bias = mean(higher_outcome_chosen, na.rm=1))
unchosen_bias_exp4 <- subset(bias_exp4, chosen_trial_name=="Unchosen") %>%
  spread(repeat_cond_name, bias) %>%
  mutate(cond_diff = Non_repeated - Repeated)

compute_power(0.80, 0.05, as.data.frame(unchosen_bias_exp4), "Repeated",0.5)
compute_power(0.80, 0.05, as.data.frame(unchosen_bias_exp4), "Non_repeated", 0.5)
compute_power(0.80, 0.05, as.data.frame(unchosen_bias_exp4), "cond_diff", 0)


```
## Decision bias analysis

### create data frames 

```{r}

# === Means ===

# compute upward bias (mean gain chosen) - for chosen and unchosen alone
upbias <- all_exps_list$final_decisions %>% 
  mutate(chosen_trial = ifelse(chosen_trial==1, "Chosen pairs", "Unchosen pairs")) %>%
  group_by(Exp, PID, chosen_trial, cond_name) %>% 
  dplyr::summarize(upbias = mean(higher_outcome_chosen, na.rm=1),
                   pair_acc = mean(pair_acc, na.rm=1),
                   choice_acc = mean(choice_acc, na.rm=1, nan.rm=1),
                   object_acc = mean(object_acc, na.rm=1, nan.rm=1))

upbias_spread <- upbias %>% spread(chosen_trial, upbias) %>%
  mutate(upbias_diff = `Chosen pairs` - `Unchosen pairs`)

# === Model fits ===
arrange_fits <- function(coef_list, Exp){
  coefs <- subset(coef_list$summary_group_coefs, grepl(c("Intercept_"),coef) | grepl(c("Slope_"),coef)) %>%
    mutate(Exp = Exp, 
           sig = ifelse((Median>0 & low95HDI>0 & high95HDI>0) | 
                          (Median<0 & low95HDI<0 & high95HDI<0),"*",""),
           value = sprintf("%.2f [%.2f %.2f]%s",Median, low95HDI, high95HDI,sig), 
           sig = ifelse((Median>0 & low95HDI>0 & high95HDI>0) | 
                          (Median<0 & low95HDI<0 & high95HDI<0),"*",""),
           value = sprintf("%.2f [%.2f %.2f]%s",Median, low95HDI, high95HDI,sig)) %>%
    separate(coef, c("coef","choice","condition"), "_")
  return(coefs)
}
fits <- bind_rows(arrange_fits(coef_list_Exp1, "Exp1"), 
                  arrange_fits(coef_list_Exp3, "Exp3"), 
                  arrange_fits(coef_list_Exp4, "Exp4"),
                  arrange_fits(coef_list_Exp5, "Exp5"),
                  arrange_fits(coef_list_Exp6, "Exp6"))

```

### find means of decision bias for chosen and unchosen pairs in Exp1 

```{r}

upbias_means_exp1 <- subset(upbias_spread, Exp=="Exp1") %>% group_by(cond_name) %>%
  dplyr::summarize(mean_chosen = mean(`Chosen pairs`, na.rm=1),
                   se_chosen = sd(`Chosen pairs`)/sqrt(n()),
                   mean_unchosen= mean(`Unchosen pairs`, na.rm=1),
                   se_unchosen = sd(`Unchosen pairs`)/sqrt(n()))
fits_exp1 <- subset(fits, Exp=="Exp1" & coef=="Intercept")

sprintf("chosen pairs (M=%.2f, SEM=%.2f, chosen intercept coefficient: %.2f [%.2f, %.2f]",
        upbias_means_exp1$mean_chosen, upbias_means_exp1$se_chosen,
        fits_exp1$Median[fits_exp1$choice=="Chosen"],fits_exp1$low95HDI[fits_exp1$choice=="Chosen"],
        fits_exp1$high95HDI[fits_exp1$choice=="Chosen"])

sprintf("unchosen pairs (M=%.2f, SEM=%.2f, unchosen intercept coefficient: %.2f [%.2f, %.2f]",
        upbias_means_exp1$mean_unchosen, upbias_means_exp1$se_unchosen,
        fits_exp1$Median[fits_exp1$choice=="Unchosen"],fits_exp1$low95HDI[fits_exp1$choice=="Unchosen"],
        fits_exp1$high95HDI[fits_exp1$choice=="Unchosen"])

```

### Analyze pairs memory and decision bias

```{r}

# === Bayesian linear regression to predict diverging bias as a function of pairs memory ===

bias_effect <- subset(all_exps_list$final_decisions, Exp=="Exp1") %>%
  group_by(PID,chosen_trial_name) %>%
  dplyr::summarize(p_gain = mean(higher_outcome_chosen, na.rm=1)) %>%
  spread(chosen_trial_name, p_gain) %>%
  mutate(bias_effect = Chosen-Unchosen)
memory_effect <- subset(all_exps_list$memory, Exp=="Exp1") %>%
  group_by(PID) %>%
  dplyr::summarise(pairs_acc = mean(pair_acc, na.rm=1))
bias_memory <- merge(bias_effect, memory_effect, by="PID")

bias_memory <- mutate(bias_memory, pairs_acc_zscore = zscore(pairs_acc), bias_zsore = zscore(bias_effect))

m_memory_exp1 <- stan_glm(data = bias_memory, bias_zsore ~ pairs_acc_zscore, family = gaussian())
summary(m_memory_exp1)
# plot
x_rng <- range(bias_memory$pairs_acc) 
x_steps <- seq(x_rng[1], x_rng[2], length.out = 96)
new_data <- data.frame(observation = seq_along(x_steps), pairs_acc = x_steps)
pred_lin <- posterior_linpred(m_memory_exp1, newdata = new_data)
tidy_predictions <- function(mat_pred, df_data, obs_name = "observation", prob_lwr = .025, prob_upr = .975) {
  # Get data-frame with one row per fitted value per posterior sample
  df_pred <- mat_pred %>% 
    as.data.frame %>% 
    setNames(seq_len(ncol(.))) %>% 
    tibble::rownames_to_column("posterior_sample") %>% 
    tidyr::gather_(obs_name, "fitted", setdiff(names(.), "posterior_sample"))
  # Helps with joining later
  class(df_pred[[obs_name]]) <- class(df_data[[obs_name]])
  # Summarise prediction interval for each observation
  df_pred %>% 
    group_by_(obs_name) %>% 
    summarise(median = median(fitted),lower = quantile(fitted, prob_lwr), upper = quantile(fitted, prob_upr)) %>%
    left_join(df_data, by = obs_name)
}
df_pred_lin <- tidy_predictions(pred_lin, new_data)

```

### Plot bias and memory for exp1

```{r}

load("../Analysis/data/Models/Memory_Bias/Coef_lists/coefs_pair_acc_bias_Exp1.RData")

# decision bias
bias <- coefs_pair_acc_bias_Exp1$bias_memory_data %>% gather(choice, bias, `0`:`1`) %>%
    mutate(choice = ifelse(choice==1, "Chosen pairs", "Unchosen pairs"), condition=NaN)
  
p1 <- ggplot(bias, aes(x=choice,y=bias)) +
        stat_summary_bin(aes(y=bias), fun.y="mean", geom="bar", binwidth=0.2, 
                         position=position_dodge(width=1), fill="grey") +
        geom_point(aes(color=factor(condition)), position=position_jitterdodge(dodge.width=1, jitter.width=0.1), 
                   fill="white", shape=21, stroke=0.4, size=3) +
        scale_color_manual(values="black") + 
        # scale_fill_manual(values=rep("black",2)) + 
        stat_summary(fun.data=mean_se, fun.args = list(mult=2), geom="errorbar", width=0.1, size=0.7, 
                     position=position_nudge(0.2), color="turquoise4") + 
        geom_hline(yintercept=0.5, size=0.5, linetype="dashed") + 
        scale_y_continuous(expand=c(0,0), breaks=c(0,0.5,1), limits=c(0,1.05)) + 
        theme + 
        theme(axis.title.x=element_blank(), legend.position="none",
              text = element_text(size=20,family="Helvetica"),
              axis.text=element_text(size=20), axis.title=element_text(size=20),
              plot.title = element_text(size=20)) +
        labs(y="p(Choose Gain)", title="Reversed preference\nfor chosen and unchosen pairs")

# memory and bias
model_text <- as.data.frame(coefs_pair_acc_bias_Exp1$posterior_draws) %>%
  dplyr::summarize(median=median(pair_acc), 
                   lower=quantile(pair_acc,0.025), 
                   upper=quantile(pair_acc,0.975)) %>%
  mutate(x=0.3, y=-0.5, text = sprintf("Slope:\n%.2f [%.2f, %.2f]",median, lower, upper))

p2 <- ggplot(coefs_pair_acc_bias_Exp1$bias_memory_data, aes(y=bias_diff,x=pair_acc)) + 
      geom_ribbon(data = mutate(coefs_pair_acc_bias_Exp1$predicted_summary_list[[1]], bias_diff=median), 
                  aes(ymin=lower, ymax=upper), fill="#DFEBE9") + 
      geom_line(aes(y=median), 
                data=coefs_pair_acc_bias_Exp1$predicted_summary_list[[1]], colour="turquoise4", size=1) +  
      geom_point(size=3, fill="white", shape=21, stroke=0.4) + 
      theme + point_plot_theme +
      geom_hline(yintercept=0, size=0.5, linetype="dashed") + 
      geom_vline(xintercept=0.5, size=0.5, linetype="dashed") +
      scale_y_continuous(expand=c(0,0), breaks=c(-1,0,1), limits=c(-1.01,1.01)) + 
      scale_x_continuous(expand=c(0,0), breaks=c(0, 0.5, 1), limits=c(-0.01, 1.01)) +
      geom_text(data=model_text, mapping=aes(x=x, y=y, label=text),size=8) + 
      theme(legend.position="none", axis.text=element_text(size=20),
            text = element_text(size=20,family="Helvetica"),
            axis.title=element_text(size=20), plot.title = element_text(size=20)) +
      labs(y="Diverging effect\n(chosen minus unchosen)",
           x="Pairs memory\n(accuracy)",
           title="Better pairs memory predicts\nmagnitude of diverging decisions bias")

p <- plot_grid(p1,p2,ncol=2,axis="bt",align="h")

ggsave(filename=sprintf("%s.%s","Figure_Exp1test",fig_type), plot=p, width=fig_size[1]+5,height=fig_size[2]-4)


```

run ttests and anova

```{r}

# ttest upbias with chance (0.5) 


```

#### Power analysis for replication study (based on Exp 2)

```{r}

library("pwr")
library("lsr")

desired_power <- 0.9
desired_sig_level <- 0.05

bias_all_exps <- all_exps_list$final_decisions %>% 
  group_by(Exp, PID, chosen_trial_name) %>%
  dplyr::summarize(upbias = mean(higher_outcome_chosen, na.rm=1))

ttest_all_exps <- bias_all_exps %>% 
  group_by(Exp, chosen_trial_name) %>%
  dplyr::summarize(t_value = as.numeric(t.test(upbias, rep(0.5, n()))["statistic"]),
            p_value_round = round(as.numeric(t.test(upbias, rep(0.5, n()))["p.value"]),3),
            p_value = as.numeric(t.test(upbias, rep(0.5, n()))["p.value"]),
            cohens_d = cohensD(x=upbias,y=rep(0.5, n()),method="paired"), # mean1 - mean2/sd(mean1)
            power = as.numeric(pwr.t.test(n = n(), d = cohens_d, sig.level = desired_sig_level, 
                               type = c("paired"))["power"]),
            desired_sample = as.numeric(pwr.t.test(power = desired_power, d = cohens_d, 
                                        sig.level = desired_sig_level, type = c("paired"))["n"]))

```


